# Data transformation

## Enhancing user records with data availability indicators

Datasets collected per user vary, due to different app version used throughout the study. Not all users have images of the meals recorded, and not all users opted-in to sharing Fitbit data (the source of activities and sleep data). Here we enhance the users records by checking on the availability of those datasets.

```{r}
user_directory_exists <- function (ids, dirname) {
  sapply(ids, function(id) dir.exists(file.path(raw_data_path, id, dirname)))
}

user_file_exists <- function(ids, filename) {
  sapply(ids, function(id) file.exists(file.path(raw_data_path, id, filename)))
}

users.all = users.raw %>%
  mutate(
    user_id = as.character(user_id),
    user_number = as.integer(user_id),
    user_fct = fct_reorder(user_id, user_number),
    app = as.factor(app),
    has_meal_images = user_directory_exists(user_id, "images"),
    has_sleep = user_file_exists(user_id, "sleep.csv"),
    has_activities_calories = user_file_exists(user_id, "activities-calories.csv"),
    has_activities_distance = user_file_exists(user_id, "activities-distance.csv"),
    has_activities_steps = user_file_exists(user_id, "activities-steps.csv")
  )
glimpse(users.all)
```

## Anonymized timestamps

All data coming out of this study was anonymized. As a part of anonymization the actual timestamps were replaced with a number of seconds since user's anchor timestamp, which is defined as the time when they took the first recording with the app. Part of the data transformation was to convert all timestamps from a 'number of seconds' format to the actual timestams (`POSIXct`) format, assuming the same anchor date for all users (`2019-01-01 00:00:00 EST`). The by-product of this transformation was the alignment of all timeseries data to the same starting point, simplifying comparison across users.

## Other field types conversions

Several fields were converted to different types to simplify further calculations and visualizations:

* `user_id` converted from numeric to character type
* `eaten_at` converted to normalized timestamp
* `kind` converted to a factor
* `premeal_bg_time` converted to normalized timestamp
* `postmeal_bg_time`converted to normalized timestamp

```{r}
as.timestamp <- function(minutes_since, anchor_date = '2019-01-01 00:00:00 EST') {
  anchor_time = as.POSIXct(anchor_date)
  return(anchor_time + minutes_since * 60)
}

as.meals_kind <- function(values) {
  return(fct_relevel(values, c("breakfast", "morning_snack", "lunch", "afternoon_snack", "dinner", "after_dinner_snack")))
}

print_df <- function(df, ...) {
  print(df %>% select(...))
  return(df)
}
```

```{r}
bg_readings.all = bg_readings.raw %>%
  mutate(
    user_id = as.character(user_id),
    user_number = as.integer(user_id),
    user_fct = fct_reorder(user_id, user_number),
    time = as.timestamp(time), 
    meals_kind = as.meals_kind(meals_kind),
    readings_kind = fct_relevel(readings_kind, c("premeal", "postmeal"))
  )
glimpse(bg_readings.all)
```

```{r}
meals.all = meals.raw %>%
  mutate(
    user_id = as.character(user_id),
    user_number = as.integer(user_id),
    user_fct = fct_reorder(user_id, user_number),
    eaten_at = as.timestamp(eaten_at),
    kind = as.meals_kind(kind),
    premeal_bg_time = as.timestamp(premeal_bg_time),
    postmeal_bg_time = as.timestamp(postmeal_bg_time)
  )
glimpse(meals.all)
```