# Missing values

## Per user data availability

Not every user appears to have every dataset available. While BG recordings and meals are recorded for all selected users, not all of them have meal images available, and not all have Fitbit data (activities and sleep) available.

```{r}
users.all %>%
  # Fake false data availability flags with NAs to visualize it as missing
  mutate(
    has_sleep = ifelse(has_sleep, has_sleep, NA),
    has_activities_calories = ifelse(has_activities_calories, has_activities_calories, NA),
    has_activities_distance = ifelse(has_activities_distance, has_activities_distance, NA),
    has_activities_steps = ifelse(has_activities_steps, has_activities_steps, NA),
  ) %>%
  visdat::vis_dat()
```

```{r}
users.all %>%
  filter(has_sleep & has_activities_steps & has_activities_calories & has_activities_distance & !is.na(complete_meals)) %>%
  count()
```

We have only 7 users with all data present (including complete meals anbd fitbit data), and that is not even taking into account the number of quality meals for those users. The consequence of this sparse data availability is that for different analysis we would need to use different slices of the overall data and different sets of users. When analyzing data for a particular user, the user selection matters a lot.


## Meals data availability

```{r fig.width=6}
visdat::vis_dat(meals.all)
```

There are several patterns of missing values that can be seen from this plot.

`platemate_id` is almost universally missing from all meals.

`premeal_bg` and `premeal_bg_time` are 100% present. The application did not allow recording a meal without also recording BG reading taken shortly before the meal. 

`postmeal_bg` and `postmeal_bg_time` on the other hand are not recorded for all meals. This is because despite of the applications sending notifications to the users 2h after the meal reminding to record a postmeal BG, users dodn't do it for every meal. Further we show a breakdown of "compliance" for taking postmeal BG, and a way to impute the missing values.

Some relatively small percentage of meal are missing the "eval" (evaluator's) fields on the nutritional meal breakdown (e.g. `calories_eval`, `eval_calories_computed`, `eval_proportion_carbs`...). 

The "user" nutritional breakdown values are mostly missing (fields like `calories_user`, `user_calories_computed`, `user_proportion_carbs`...). 

```{r fig.height=3}
extracat::visna(meals.raw)
```

## Visualizing user compliance on recording postmeal BG

```{r fig.width=3,fig.asp=1.5}
meals.all %>%
  mutate(premeal_bg_exist = ifelse(is.na(premeal_bg), 0, 1), postmeal_bg_exist = ifelse(is.na(postmeal_bg), 0, 1)) %>%
  group_by(user_id) %>%
  summarise(all_meals = sum(premeal_bg_exist), Present = sum(postmeal_bg_exist)) %>%
  ungroup() %>%
  mutate(Missing = all_meals - Present) %>%
  arrange(desc(all_meals)) %>%
  gather(key = "type", value = "Freq", -all_meals, -user_id) %>%
  ggplot() +
  geom_col(aes(x = fct_reorder(user_id, all_meals), y = Freq, fill = fct_relevel(type, "Present", "Missing")), position = position_stack(reverse = TRUE)) +
  coord_flip() + 
  labs(
    title = "Meals with missing postmeal BG by user",
    x = "Users",
    y = "Counts",
    fill = "Postmeal BG"
  ) +
  theme(legend.position = "right")
```

There does not seem to be any patterns in this visualization beyond staing that some users were diligent and recorded most of the postmeals BG, some totally ignored it, and the majority of users are somewhere in between.

## Imputing post-meal BG

As shown before, a lot of meals are missing postmeal BG records, indicating that the users did not record another BG reading around 2h mark as instructed. However, the value of postmeal_bg and postmeal_bg_time can be imputed, considering that in some cases users ate the next meal and recorded a BG reading with it. We impute missing postmeal values with the next meal premeal values. 




```{r}
meals.imputed = meals.all %>%
  arrange(user_id, eaten_at) %>%
  mutate(
    postmeal_bg_time = coalesce(postmeal_bg_time, lead(premeal_bg_time)), 
    postmeal_bg = coalesce(postmeal_bg, lead(premeal_bg))
  ) %>%
  mutate(
    premeal_bg_delay_minutes = as.double(difftime(eaten_at, premeal_bg_time, units = "mins")),
    postmeal_bg_delay_minutes = as.double(difftime(postmeal_bg_time, eaten_at, units = "mins")),
    bg_impact = postmeal_bg - premeal_bg,
    bg_impact_slope = bg_impact / (premeal_bg_delay_minutes + postmeal_bg_delay_minutes)
  ) %>%
  # keep only positive delays, this filters out boundary meals when user_id changes in the dataset and getting the next record meal will be incorrect.
  filter(!is.na(premeal_bg_delay_minutes) & premeal_bg_delay_minutes >= 0 & !is.na(postmeal_bg_delay_minutes) & postmeal_bg_delay_minutes >= 0 ) 
```

After imputing `postmeal_bg` and `postmeal_bg_time` several new metrics were calculated:

* `premeal_bg_delay_minutes` calculated number of minutes before premeal BG is taken and the meal
* `postmeal_bg_delay_minutes` calculated number of minutes after the meal when postmeal BG was taken
* `bg_impact` calculated difference between `postmeal_bg` and `premeal_bg`
* `bg_impact_slope` ratio of `bg_impact` over the duration in minutes between post-meal and pre-meal times. Higher `bg_impact_slope` means faster BG increase abter the meal



```{r fig.width=6}
visdat::vis_dat(meals.imputed)
```

## Analyzing pre and post meal BG time intervals 

THere were different versions of the app. Some versions of the app asked users to enter nutritional estimation, but some version only asked on the goals, but not nutritional.


```{r fig.height=2}
g1 = meals.imputed %>%
  #filter(user_id %in% users.selected[["user_id"]]) %>%
  filter(premeal_bg_delay_minutes >= 0 & premeal_bg_delay_minutes <= 60) %>%
  ggplot() +
  geom_boxplot(aes(y = -premeal_bg_delay_minutes)) + 
  scale_y_continuous(breaks=c(0, -15, -30, -45, -60), labels = c("0", "15m", "30m", "45m", "1h"), limits = c(-60, 0), expand = expand_scale(mult = c(0.1, 0))) + 
  coord_flip() +
  labs( 
    y = "Premeal delay"
  ) + 
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

g2 = meals.imputed %>%
  filter(user_id %in% users.selected[["user_id"]]) %>%
  filter(postmeal_bg_delay_minutes >= 0 & postmeal_bg_delay_minutes <= 1000) %>%
  ggplot() +
  geom_boxplot(aes( y = postmeal_bg_delay_minutes), varwidth = TRUE) + 
  scale_y_continuous(breaks=c(0, 60, 120, 180, 240, 300), labels = c("0", "1h", "2h", "3h", "4h", "5h"), limits = c(0, 1000), expand = expand_scale(mult = c(0, .05))) + 
  coord_flip() +
  labs( 
    y = "Postmeal delay"
  ) + 
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

gridExtra::grid.arrange(
  g1, g2, nrow = 2,
  heights = c(1, 1),
  top = "Pre and Post meal delays for BG"
)

```

Talking point: at some version the users were prevented from entering post meal before they are entered premeal, and users were forced to re-enter (backfill) previous meal BG. This led to a lot og outlyers.


TODO: explain how the data was backfilled, not every user captured post BG in the app, even after reminders. However later on the association between the meal and pre and post BG were re-calculated based on the timestamps and pre/post BG were picked as timestamps around the meal.

### Selecting users for the analysis

Not all users used the application in the same capacity, and some recorded very few meals. For the further analysis we select only those users that recorded at least 14 meals (threshold picked such that 28 users were selected for easy facetting in a grid)

```{r}
selected_user_ids = meals.imputed  %>%
  filter(premeal_bg_delay_minutes >= 0 & premeal_bg_delay_minutes <= 60) %>% # Limit premeal bg delay to 1h
  filter(postmeal_bg_delay_minutes >= 0 & postmeal_bg_delay_minutes <= 240) %>% # Limit postmeal bg delay to 4h 
  group_by(user_id) %>%
  summarise(Freq = n()) %>%
  ungroup() %>%
  arrange(desc(Freq)) %>%
  top_n(24, Freq)
selected_user_ids = selected_user_ids[['user_id']]
print(selected_user_ids)
```

## Final clean meals dataset

Limit the final meals dataset with filtered users

```{r}
meals.selected = meals.imputed %>%
  filter(user_id %in% selected_user_ids) %>%
  filter(premeal_bg_delay_minutes >= 0 & premeal_bg_delay_minutes <= 60) %>% # Limit premeal bg delay to 1h
  filter(postmeal_bg_delay_minutes >= 0 & postmeal_bg_delay_minutes <= 240) # Limit postmeal bg delay to 4h 
```



