# Missing values

## Per user data availability

Not every user appears to have every dataset available. While BG recordings and meals are recorded for all selected users, not all of them have meal images available, and not all have Fitbit data (activities and sleep) available.

```{r}
users %>%
  # Fake false values with NAs to visualize it as missing
  mutate(
    has_sleep = ifelse(has_sleep, has_sleep, NA),
    has_activities_calories = ifelse(has_activities_calories, has_activities_calories, NA),
    has_activities_distance = ifelse(has_activities_distance, has_activities_distance, NA),
    has_activities_steps = ifelse(has_activities_steps, has_activities_steps, NA),
  ) %>%
  visdat::vis_dat()
```

## Meals data availability

```{r fig.width=6}
visdat::vis_dat(meals_data)
```

```{r fig.height=3}
extracat::visna(meals_data)
```


### Selecting users for the analysis

Not all users used the application in the same capacity, and some recorded very few meals. For the further analysis we select only those users that recorded at least 14 meals (threshold picked such that 28 users were selected for easy facetting in a grid)

```{r}
selected_users = users %>%
  filter(complete_meals > 14)
print(selected_users[["user_id"]])
```

Limit the final meals dataset with filtered users

```{r}
meals_data = meals_data %>% filter(user_id %in% selected_users[["user_id"]])
```

## Missing values in meals_data


THere were different versions of the app. Some versions of the app asked users to enter nutritional estimation, but some version only asked on the goals, but not nutritional.


```{r fig.height=12, fig.width=6}
meals_data %>%
  filter(premeal_bg_delay_minutes > 0 & premeal_bg_delay_minutes < 60) %>%
  ggplot() +
  geom_boxplot(aes(x = user_id, y = premeal_bg_delay_minutes), varwidth = TRUE) + 
  scale_y_continuous(breaks=c(0, 15, 30, 45, 60), labels = c("0", "15min", "30min", "45min", "1h"), limits = c(0, 60)) + 
  coord_flip() +
  labs( title = "" )
```

Talking point: at some version the users were prevented from entering post meal before they are entered premeal, and users were forced to re-enter (backfill) previous meal BG. This led to a lot og outlyers.


```{r fig.height=12, fig.width=6}
meals_data %>%
  ggplot() +
  geom_boxplot(aes(x = user_id, y = postmeal_bg_delay_minutes), varwidth = TRUE) + 
  scale_y_continuous(breaks=c(0, 60, 120, 180, 240), labels = c("0", "1h", "2h", "3h", "4h"), limits = c(0, 1000)) + 
  coord_flip() +
  labs( title = "" )
```

TODO: explain how the data was backfilled, not every user captured post BG in the app, even after reminders. However later on the association between the meal and pre and post BG were re-calculated based on the timestamps and pre/post BG were picked as timestamps around the meal.


## Imputing post-meal BG

```{r}

```