# Missing values

## Per user data availability

Not every user appears to have every dataset available. While BG recordings and meals are recorded for all selected users, not all of them have meal images available, and not all have Fitbit data (activities and sleep) available.

```{r}
users.all %>%
  # Fake false data availability flags with NAs to visualize it as missing
  mutate(
    has_sleep = ifelse(has_sleep, has_sleep, NA),
    has_activities_calories = ifelse(has_activities_calories, has_activities_calories, NA),
    has_activities_distance = ifelse(has_activities_distance, has_activities_distance, NA),
    has_activities_steps = ifelse(has_activities_steps, has_activities_steps, NA),
  ) %>%
  visdat::vis_dat()
```

```{r}
users.all %>%
  filter(has_sleep & has_activities_steps & has_activities_calories & has_activities_distance & !is.na(complete_meals)) %>%
  count()
```

We have only 7 users with all data present (including complete meals anbd fitbit data), and that is not even taking into account the number of quality meals for those users. The consequence of this sparse data availability is that for different analysis we would need to use different slices of the overall data and different sets of users. When analyzing data for a particular user, the user selection matters a lot.


## Meals data availability

```{r}
visdat::vis_dat(meals.all)
```

There are several patterns of missing values that can be seen from this plot.

`platemate_id` is almost universally missing from all meals.

`premeal_bg` and `premeal_bg_time` are 100% present. The application did not allow recording a meal without also recording BG reading taken shortly before the meal. 

`postmeal_bg` and `postmeal_bg_time` on the other hand are not recorded for all meals. This is because despite of the applications sending notifications to the users 2h after the meal reminding to record a postmeal BG, users dodn't do it for every meal. Further we show a breakdown of "compliance" for taking postmeal BG, and a way to impute the missing values.

Some relatively small percentage of meal are missing the "eval" (evaluator's) fields on the nutritional meal breakdown (e.g. `calories_eval`, `eval_calories_computed`, `eval_proportion_carbs`...). 

The "user" nutritional breakdown values are mostly missing (fields like `calories_user`, `user_calories_computed`, `user_proportion_carbs`...). The reason is that some versions of the app asked users to enter nutritional estimation, but some version only asked on the goals, but not nutritional estimation.


```{r}
extracat::visna(meals.all)
```

## Visualizing user compliance on recording postmeal BG

```{r fig.asp=1.5}
meals.all %>%
  mutate(premeal_bg_exist = ifelse(is.na(premeal_bg), 0, 1), postmeal_bg_exist = ifelse(is.na(postmeal_bg), 0, 1)) %>%
  group_by(user_id) %>%
  summarise(all_meals = sum(premeal_bg_exist), Present = sum(postmeal_bg_exist)) %>%
  ungroup() %>%
  mutate(Missing = all_meals - Present) %>%
  arrange(desc(all_meals)) %>%
  gather(key = "type", value = "Freq", -all_meals, -user_id) %>%
  ggplot() +
  geom_col(aes(x = fct_reorder(user_id, all_meals), y = Freq, fill = fct_relevel(type, "Present", "Missing")), position = position_stack(reverse = TRUE)) +
  coord_flip() + 
  labs(
    title = "Meals with missing postmeal BG by user",
    x = "Users",
    y = "Counts",
    fill = "Postmeal BG"
  ) +
  theme(legend.position = "right")
```

There does not seem to be any patterns in this visualization beyond staing that some users were diligent and recorded most of the postmeals BG, some totally ignored it, and the majority of users are somewhere in between.

## Imputing post-meal BG

As shown before, a lot of meals are missing postmeal BG records, indicating that the users did not record another BG reading around 2h mark as instructed. However, the value of postmeal_bg and postmeal_bg_time can be imputed, considering that in some cases users ate the next meal and recorded a BG reading with it. We impute missing postmeal values with the next meal premeal values. 

```{r}
meals.imputed = meals.all %>%
  arrange(user_id, eaten_at) %>%
  mutate(
    postmeal_bg_time = coalesce(postmeal_bg_time, lead(premeal_bg_time)), 
    postmeal_bg = coalesce(postmeal_bg, lead(premeal_bg))
  ) %>%
  mutate(
    premeal_bg_delay_minutes = as.double(difftime(eaten_at, premeal_bg_time, units = "mins")),
    postmeal_bg_delay_minutes = as.double(difftime(postmeal_bg_time, eaten_at, units = "mins")),
    bg_impact = postmeal_bg - premeal_bg,
    bg_impact_slope = bg_impact / (premeal_bg_delay_minutes + postmeal_bg_delay_minutes)
  ) %>%
  # keep only positive delays, this filters out boundary meals when user_id changes in the dataset and getting the next record meal will be incorrect.
  filter(!is.na(premeal_bg_delay_minutes) & premeal_bg_delay_minutes > 0 & !is.na(postmeal_bg_delay_minutes) & postmeal_bg_delay_minutes > 0 ) 
```

After imputing `postmeal_bg` and `postmeal_bg_time` several new metrics were calculated:

* `premeal_bg_delay_minutes` calculated number of minutes before premeal BG is taken and the meal
* `postmeal_bg_delay_minutes` calculated number of minutes after the meal when postmeal BG was taken
* `bg_impact` calculated difference between `postmeal_bg` and `premeal_bg`
* `bg_impact_slope` ratio of `bg_impact` over the duration in minutes between post-meal and pre-meal times. Higher `bg_impact_slope` means faster BG increase abter the meal

## Analyzing pre and post meal BG time intervals 

Here we look at the distribution of values for premeal BG delay (defined as the difference in minutes between users recording a BG reading and eating a meal), and postmeal BG delay (defined as the difference in minutes between users eating a meal and taking a postmeal BG reading).

```{r}
g1 = meals.imputed %>%
  filter(premeal_bg_delay_minutes >= 0 & premeal_bg_delay_minutes <= 60) %>%
  ggplot() +
  geom_histogram(aes(x = premeal_bg_delay_minutes), binwidth = 1.0) +
  scale_x_continuous(breaks=c(0, 15, 30, 45, 60), labels = c("0", "15m", "30m", "45m", "1h"), limits = c(0, 60), expand = expand_scale(mult = c(0, 0.1))) + 
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

g2 = meals.imputed %>%
  filter(premeal_bg_delay_minutes >= 0 & premeal_bg_delay_minutes <= 60) %>%
  ggplot() +
  geom_boxplot(aes(y = premeal_bg_delay_minutes)) + 
  scale_y_continuous(breaks=c(0, 15, 30, 45, 60), labels = c("0", "15m", "30m", "45m", "1h"), limits = c(0, 60), expand = expand_scale(mult = c(0, 0.1))) + 
  coord_flip() +
  labs( 
    y = "Premeal delay"
  ) + 
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


g3 = meals.imputed %>%
  filter(postmeal_bg_delay_minutes >= 0 & postmeal_bg_delay_minutes <= 1000) %>%
  ggplot() +
  geom_histogram(aes(x = postmeal_bg_delay_minutes), binwidth = 30) +
  scale_x_continuous(breaks=c(0, 60, 120, 180, 240, 300), labels = c("0", "1h", "2h", "3h", "4h", "5h"), limits = c(0, 1000), expand = expand_scale(mult = c(0, .05))) + 
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

g4 = meals.imputed %>%
  filter(postmeal_bg_delay_minutes >= 0 & postmeal_bg_delay_minutes <= 1000) %>%
  ggplot() +
  geom_boxplot(aes(y = postmeal_bg_delay_minutes), varwidth = TRUE) + 
  scale_y_continuous(breaks=c(0, 60, 120, 180, 240, 300), labels = c("0", "1h", "2h", "3h", "4h", "5h"), limits = c(0, 1000), expand = expand_scale(mult = c(0, .05))) + 
  coord_flip() +
  labs( 
    y = "Postmeal delay"
  ) + 
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

gridExtra::grid.arrange(
  g1, g2, g3, g4, nrow = 4,
  heights = c(3, 1, 3, 1),
  top = "Pre and post meal BG delays"
)

```

Premeal BG delay distribution has two distinct modes, one at around 5 min and another at around 22 min. The application to record meals and premeal BGs asked the users to record those two pieces of information at the same time, but also allowed to change the time when the actual BG test was performed. The explanation for the first mode could be the interaction scenario when a person took a BG reading, and before the meal has started recorded it with the application (adjusting the BG test time to about 5min back). The socond mode could be another interaction scenario, where a person takes a BG reading, then eats the meal, and only after the meal is finished records the meal in the app (adjusting the BG test time to about 22 min back). One way to co nfirm it would be to visually analyze the meal images for the meals belonging to the second mode and see if those are the images of already completed meals. The researcher confirmed that in focus groups some people found the process of recording a meal to be disruptive, and may be we are seeing a hack some users adopted not to disrupt the meal experience with a lengthy process of entering a meal in the app while they are hungry.

The postmeal BG delays have a median around 2h 15min, due the fact that the app was sending reminders at around 2h after the meal. The distribution has a long tail and a lot of outliers, probably causes by our postmeal BG imputation method and/or lapses in consistently using the app for all meals. For any analysis ths focuses on the question of glycemic impact of the meals such long postmeal interval in not useful. In our analysis we filtered the meal to only those that have a postmeal BG at least 4h after the meal.

### Selecting users for the analysis

Not all users used the application in the same capacity, and some recorded very few meals. For the further analysis we select only those users that recorded at least 14 meals (threshold picked such that 28 users were selected for easy facetting in a grid)

```{r}
users.selected = meals.imputed  %>%
  filter(premeal_bg_delay_minutes > 0 & premeal_bg_delay_minutes <= 60) %>% # Limit premeal bg delay to 1h
  filter(postmeal_bg_delay_minutes > 0 & postmeal_bg_delay_minutes <= 240) %>% # Limit postmeal bg delay to 4h 
  group_by(user_id) %>%
  summarise(Freq = n()) %>%
  ungroup() %>%
  arrange(desc(Freq)) %>%
  head(24) # 24 chosed to fit in a 6x4 facet grid
selected_user_ids = users.selected[['user_id']]
print(selected_user_ids)
```

## Final selected meals subset for BG impact analysis

We limit the list of all meals with imputed values to only those users with the most meals and where pre and post meal delays are withing acceptable range (1h for premeal delay and 4h for postmeal delay).

```{r}
meals.selected = meals.imputed %>%
  filter(user_id %in% selected_user_ids) %>%
  filter(premeal_bg_delay_minutes > 0 & premeal_bg_delay_minutes <= 60) %>% # Limit premeal bg delay to 1h
  filter(postmeal_bg_delay_minutes > 0 & postmeal_bg_delay_minutes <= 240) # Limit postmeal bg delay to 4h 
```



