# Results

Below we include some of the visualizations selected to highlight interesting insights about the data. In our exploratory analysis, we have tried many different types of visualizations. However, not all visualizations we tried were insightful or revealed any interesting patterns; these visualizations are not included in the main results section. Some of those plots are included in the appendix section.

We heavily utilized facets to highlight patterns and differences between users. These faceted plots provide only a rough overview of an individual user’s stats, and mainly focus on the contrast and comparison between users. In order to facilitate comparison of the same users across different plots we maintained consistent placement of users on the facet grid, so that the same user consistently appears in the same position on all faceted plots.

One of the discoveries in this project was that aggregating data across users does not yield many interesting insights. This could be due to the relatively small number of users in the dataset. However, this could also be due to the high individual variability in both behaviors (nutritional composition of meals) and biomarkers (BG before and after meals). This variability suggests the need to examine patterns for each user and compare patterns across users, rather than aggregating data, thus supporting our approach to using faceted visualizations.

## Looking at the time of the meals

```{r}
meals.selected %>%
  mutate(eaten_at_hour = hour(eaten_at)) %>%
  ggplot() +
  geom_histogram(aes(x=eaten_at_hour), bins = 24, fill = "#4285FF") +
  facet_wrap(~user_fct, ncol = 6) + 
  labs(
    title = "Time of day when users had a meal",
    x = "Hour of day",
    y = "Number of meals"
  ) + 
  theme(
    #strip.text = element_text(size=25)
  )
```

Examination of time of meals across different users shows different eating patterns between individuals. Some users (like 2004, 2262, 2392, 24, 2475, 2890, 56, 57, and 58) have a well-defined 3 humps pattern. These users tend to eat 3 meals per day at regular times. Other users (like 2942, 2948, and 2800) have a defined 2 hump pattern; these individuals tend to eat only 2 meals a day. Yet other users (like 2254, 2536) do not have clear humps in their meal patterns; these users exhibit behaviors that are typically described as “grazing” or eating at any time of day with no consistent times. 

For those “two meals” user, the question remains as to what meal types (breakfast, lunch, etc.) these users tend to skip? Below we analyze meal types per user.

## Meal type analysis

```{r}
meals.selected %>%
  group_by(user_fct, kind) %>%
  summarise(freq = n()) %>%
  group_by(user_fct) %>%
  mutate(prop = freq/sum(freq)) %>%
  ungroup() %>%
  ggplot() +
  geom_bar(aes(x = kind, y = prop, fill = kind), stat = "identity") + 
  facet_wrap(~user_fct, ncol = 6) + 
  labs(
    x = "Kind of meal",
    y = "Proportion",
    title = "Proportion of Meals"
  ) + 
  guides(fill = guide_legend(nrow = 1)) +
  theme(
    legend.title = element_blank(), 
    legend.position="bottom",
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
```

This analysis showed both consistency and variability across users. 

On one hand, in this dataset, breakfast is prominently featured for almost all users and appears to be the most consistent meal across all individuals.

On the other hand, there is much variability in regards to the other daily meals. For example, user 2254 (classified as “grazer”) does indeed eat a lot of snacks, which is consistent with the distribution of their times of meals. Some of the typical “two meals” users (i.e., 2942 and 2800) include breakfasts and lunches, but not dinners. Yet other “two-meal” users (like 2948) do not follow the same pattern. Ther meal times are clearly centered around just two key times, but meal type distribution contains breakfast, lunch and dinner.

Another observation is that some users eat a lot of snacks, and some users do not.

## Meal nutrients analysis

```{r}
meals.selected %>%
  filter(!is.na(calories_eval)) %>%
  select(user_fct, kind, Carbs = carbs_eval, Protein = protein_eval, Fat = fat_eval, Fiber = fiber_eval) %>%
  gather(key = "Nutrient", value, -user_fct, -kind) %>%
  group_by(user_fct, Nutrient) %>%
  summarise(total=sum(value)) %>%
  group_by(user_fct) %>%
  mutate(prop = total/sum(total)) %>%
  ungroup() %>%
  mutate(Nutrient = fct_relevel(Nutrient, "Carbs", "Protein", "Fat", "Fiber")) %>% 
  ggplot() +
  geom_bar(aes(x = Nutrient, y = prop, fill = Nutrient), stat = "identity") + 
  facet_wrap(~user_fct, ncol = 6) + 
  labs(
    x = "Nutrient",
    y = "Proportion",
    title = "Proportion of nutrients across all meals"
  ) + 
  guides(fill = guide_legend(nrow = 1)) +
  theme(
    legend.title = element_blank(), 
    legend.position="bottom",
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
  
```

Of note in this analysis, carbohydrates are the biggest contributor to most users’ diet; the amount of carbohydrates in users’ meals exceed amounts of protein, then fat and then fiber. Users that support this trend are 57, 1884, 2948.

## Meal BG impact visualization

In this plot, each line segment represents a meal captured by a user; the color of the line indicates the type of the meal (breakfast, dinner, etc.). The line segment starts at the time (relative to the meal time) and BG of pre-meal measurement, and ends at the time/BG of the post-meal measurement. The steeper slope of the line indicates higher BG impact of a meal. The red horizontal line shows the diagnostic threshold level of 200mg/dL for post-meal BG to diagnose diabetes.

```{r fig.asp=0.8}
meals.selected %>% 
  ggplot() +
  geom_segment(aes(x = -premeal_bg_delay_minutes, xend = postmeal_bg_delay_minutes, y=premeal_bg, yend=postmeal_bg, color = kind), show.legend = FALSE, alpha = 0.2) + 
  # draw rect purely so that the legend appears as a solid rect, for a line with low alpha which makes it hard to see colors.
  # draw the rects outside of visible X limits set for the X axis.
  geom_rect(aes(xmin = -1000-premeal_bg_delay_minutes, xmax = -1000+postmeal_bg_delay_minutes, ymin=premeal_bg, ymax=postmeal_bg, fill = kind)) + 
  geom_hline(aes(yintercept=200), color="red", show.legend = FALSE, alpha = 0.5) +
  scale_x_continuous(breaks=c(-30, 0, 60, 120, 180, 240), labels = c("", "0", "1h", "2h", "3h", "4h"), limits = c(-30, 240)) + 
  facet_wrap(~user_fct, ncol = 6) + 
  labs(
    x = "Time since meal",
    y = "BG",
    title = "BG impact of meals per user",
    caption = "Red horizontal line indicates warning postmeal BG levele of 200mg/dL"
  ) + 
  guides(color = guide_legend(nrow = 1), fill = guide_legend(nrow = 1)) +
  theme(
    legend.title = element_blank(), 
    legend.position="bottom",
  )
```

As this plot suggests, most of the users in dataset have relatively stable blood glucose levels with relatively little change in BG from pre-meal to post-meal; few BG levels in this dataset across users exceed the warning threshold of 200 mg/dL. However, some users (namely 2254, 1809, 2262) have considerable variability in their blood glucose levels with BG trends post-meal split between donward (descreese in BG from pre-meal to post-meal) and upward (increase in BG from pre-meal to post-meal). One user in this dataset (1809) has BG levels that are consistently above the 200 level all the time.

There are also notable differences in how individuals’ BG levels change through the days. For example, user 2312 appears to be consistently starting their days with very low BG (in the 50-60 range), with their BG moving to the 100-150 range post-breafast. This behavior and impact are quite consistent across several users.

User 1821 recorded many dinners (also visible on “Proportion of meals” plot); these dinners appear to have a relatively mild impact on their BG, with a bump of around 50 BG points per dinner.

User 2942 appears to consistently have breakfasts with an impact of about 25-30 BG points and then lunch that seems to actually lower the BG in postmeal recording. That may indicate that the user takes insulin to manage diabetes.

## Trends in calories

```{r}
meals.selected %>%
  mutate(eaten_at_date = as.Date(eaten_at, tz = "EST")) %>%
  filter(eaten_at_date <"2019-04-01" & !is.na(calories_eval)) %>%
  group_by(user_fct, eaten_at_date) %>%
  summarise(sum_by_day = sum(eval_calories_computed)) %>%
  ggplot(aes(x=eaten_at_date, y=sum_by_day))+
  geom_point(size=0.1) +
  geom_smooth(method = "lm", na.rm=TRUE, size=0.5) +
  facet_wrap(~user_fct, ncol=6) +
  labs (
    title = "Trend of daily calories consumed by user",
    x = "Time",
    y = "Calories"
  ) + 
  theme(
    axis.text.x = element_text(angle = 90)
  )
```

At a glance it appears that most users either maintained or decreased the number of calories they consume per day. Notable exception is user 2890, whose calories consumption increased. One notable observation in this plot is that for most users the daily calories consumption is well below 2000 calories/day typically recommended for an average individual. One explanation is that users didn’t record all meals in a day.

Plotting of daily amount of carbs consumed by users (carbs is the most important nutrient to manage for diabetics) shows similar trends as calories; these plots are included in the appendix.

## Trend in fasting BG

```{r}
meals.selected %>%
  mutate(eaten_at_date = as.Date(eaten_at, tz = "EST")) %>%
  filter(eaten_at_date <"2019-04-01" & !is.na(premeal_bg)) %>%
  filter(kind == 'breakfast') %>%
  group_by(user_fct, eaten_at_date) %>%
  summarise(fasting_bg = first(premeal_bg)) %>%
  ggplot(aes(x=eaten_at_date, y=fasting_bg))+
  geom_point(size=0.1) +
  geom_smooth(method = "lm", na.rm=TRUE, size=0.5) +
  geom_hline(aes(yintercept=125), color="red", show.legend = FALSE, alpha = 0.5) +
  facet_wrap(~user_fct, ncol=6) +
  labs (
    title = "Trend of fasting BG by user",
    x = "Time",
    y = "Fasting BG"
  ) + 
  theme(
    axis.text.x = element_text(angle = 90)
  )
```

Fasting BG is defined as the first BG measurement after 8h of fasting, normally after sleep in the morning before breakfast. This plot shows the trend of fasting BG by user with a trend line (in blue) and a threshold (in red). Clinical guidelines in regards to the fasting BG are as follows: below 100 is considered normal (not shown on the plot), 100-125 is pre-diabetes, above 125 leads to a diagnosis of diabetes.

For most of the users the trend line looks flat and at or below the threshold level. User 1809 shows a large drop in fasting BG, which would be promising, but unfortunately only few (5) datapoint exist for that user.

## Meal images 

Here we explore the images of meals for user 2254. The first set of images shows meals that have the biggest impact over shortest timeframe (bg_impact_slope), and the second set shows the meals that had smallest (even negative) impact on BG.

```{r fig.asp=0.8}
display_meals <- function(df, ncol = 6, nrow = 3, rotate = 0) {
  meals = df[1:(ncol*nrow), ] %>% select(user_id, meal_id)
  filenames = paste0(normalizePath("~/Dropbox/MealyzerData"), "/", meals[["user_id"]], "/images/medium/", meals[["meal_id"]], ".jpg")
  data_uris = sapply(filenames, knitr::image_uri)
  images = tibble(src = data_uris)
  r2d3(data=images, script = "display_meals.js", container = "div", options=list(ncol=ncol, nrow=nrow, rotate=rotate))
}

meals.selected %>%
  filter(user_id == "2254") %>%
  arrange(desc(bg_impact_slope)) %>%
  display_meals(ncol = 6, nrow = 3, rotate = -90)
```


```{r fig.asp=0.8}
meals.selected %>%
  filter(user_id == "2254") %>%
  arrange(bg_impact_slope) %>%
  display_meals(ncol = 6, nrow = 3, rotate = -90)
```

Unsurprisinly the lowest BG impact images are all pictures of water/drinks.


### Per user correlation between variables


```{r fig.asp=1,fig.width=12}
meals.selected %>%
  filter(user_id == "2254") %>%
  filter(fiber_eval < 20) %>% # filter outliers
  filter(bg_impact_slope < 5 & bg_impact_slope > -3) %>%
  mutate(
    protein_fat=eval_proportion_protein-eval_proportion_fat, 
    protein_carb=eval_proportion_protein-eval_proportion_carbs, 
    carb_fat=eval_proportion_carbs-eval_proportion_fat) %>%
  select(calories_eval, eval_calories_computed, carbs_eval, fat_eval, protein_eval, fiber_eval, eval_proportion_carbs, eval_proportion_fat, eval_proportion_protein, bg_impact, bg_impact_slope, premeal_bg, postmeal_bg) %>%
  lattice::splom()
```

Examination of correlation between different variables per user revealed several notable trends.

First, there is a very strong correlation between `calories_eval` (calories assigned by the evaluator) and `eval_calories_computed` (calories calculated from the nutritional breakdown of a meal into carbs/fat/protein amounts). This correlation shows that evaluator’s experience is spot on estimating the meal or the evaluators are using the same formula themselves.

Second, there is a positive correlation between `calories_eval`/`eval_calories_computed` and each of `carbs_eval`, `fat_eval` and `protein_eval`. Again, no surprise here, as calories in meals can be derived from nutrients.

Next, a strong negative correlation between `eval_proportion_carbs` and `eval_proportion_fat` shows that individuals have several typical meal patterns: they either eat meals that are high in carbohydrate and low in fat, or meals that are low in carbohydrates and high in fat. Interestingly, the proportion of protein stays somewhat independent.

The correlation between `bg_impact` and `bg_impact_slope`  makes sense as the time delay between premeal and postmeal BG is relatively constant at about 2h.

Unfortunately, there are no clear correlations to suggest what factors affects the glycemic impact of meals. This relationship is non-linear; more sophisticated methods should be employed to answer this question.

## Looking into meal nutrients estimation (evaluator vs user)

Here we examine the following question: how well do users estimate nutritional composition of their own meals by comparing users’ estimates to estimates for the same meals generated by the professional nutritionist evaluators. Here, we display three plots with increasing level of details.

```{r fig.asp=2}
(function(meals) {
  estimate_diffs = meals %>%
    mutate(
      Calories = calories_user - calories_eval,
      Carbs = carbs_user - carbs_eval,
      Protein = protein_user - protein_eval,
      Fat = fat_user - fat_eval,
      Fiber = fiber_user - fiber_eval,
    ) %>%
    filter(!is.na(Calories) & !is.na(Carbs) & !is.na(Protein) & !is.na(Fat) & !is.na(Fiber))

  g1 = estimate_diffs %>%
    ggplot() +
    geom_boxplot(aes(y = Calories)) +
    geom_hline(aes(yintercept=0), color="black", alpha = 0.5) +
    scale_y_continuous(limits = c(-2000, 2000)) +
    coord_flip() + 
    labs(
      title = "Calories estimation error overall",
      y = "Error (calories)"
    ) + theme(
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    ) # + 
    #cowplot::theme_cowplot(font_size = 10)

  g2 = estimate_diffs %>%
    gather(key = "Nutrient", value = "Error", Carbs, Protein, Fat, Fiber) %>%
    ggplot() +
    geom_boxplot(aes(x = reorder(Nutrient, Error, median), y = Error)) +
    geom_hline(aes(yintercept=0), color="black", alpha = 0.5) +
    scale_y_continuous(limits = c(-100, 100)) +
    coord_flip() + 
    labs(
      title = "Nutrients estimation error overall",
      x = "Nutrient",
      y = "Error (grams)"
    )# + 
    #cowplot::theme_cowplot(font_size = 10)
  
  g3 = estimate_diffs %>%
    filter(!(user_id %in% c("99", "124", "377", "445", "1663"))) %>%
    ggplot() +
    geom_boxplot(aes(x = fct_rev(user_fct), y = Carbs), varwidth = TRUE) +
    geom_hline(aes(yintercept=0), color="black", alpha = 0.5) +
    scale_y_continuous(limits = c(-100, 100)) +
    coord_flip() + 
    labs(
      title = "Carbs estimation error by user",
      x = "Users",
      y = "Error (grams)"
    )# +
    #cowplot::theme_cowplot(font_size = 10)

    cowplot::plot_grid(g1, g2, g3, align = "hv", nrow = 3, axis = "lr", rel_heights = c(0.8, 2, 5))
})(meals.imputed)
```


The first plot displays the distribution of errors by all users when estimating meal calories. By error we mean the difference between evaluator’s estimate of the amount of calories and the user’s estimate of the same. The distribution of error looks to be a compact symmetric distribution with a median very close to 0. In the task of calorie estimation, the users achieved high accuracy as compared to professional estimation. This was a surprising finding to the researcher, who assumed before seeing the data that there will be high discrepancy between users’ estimates and professionals’ estimates, reflection low levels of nutritional knowledge among users. One consequence of this discovery is that there appears to be less need to obtain explensive professional nutritional evaluations as the users’ own estimates are comparable to those of professional evaluators.  

The second plot displays how well users estimated each nutrient overall. These plots show that on average, individuals tend to overestimate the amount of fiber in their meals, which may suggest an overly optimistic view as individuals with diabetes are typically recommended to increase fiber in their meals.  For Fat and Protein the median of the estimates are pretty close to 0. Finally, for carbohydrates, users estimates are slightly below those of the professional nutritionists. This find presents a concern as carbohydrates in meals are the main contributor to rise on post-meal BG levels; underestimating carbohydrates in meals can lead to overconsumption and to undesired increase in BG levels.

The third plot breaks down the user’s ability to estimate carbs by each user. Here we see that there are systematic biases in how different users estimate carbs. Some users (e.g. the first five users) estimate very close to the evaluator’s value, and some users are consistently under-estimating. Interestingly there are no users who systematically over-estimated the amount of carbs.
