# Appendix


```{r}
meals.selected %>%
  mutate(eaten_at_date = as.Date(eaten_at, tz = "EST")) %>%
  filter(eaten_at_date <"2019-04-01" & !is.na(carbs_eval)) %>%
  group_by(user_fct, eaten_at_date) %>%
  summarise(sum_by_day = sum(carbs_eval)) %>%
  ggplot(aes(x=eaten_at_date, y=sum_by_day))+
  geom_point(size=0.1)+
  geom_smooth(method = "lm")+
  facet_wrap(~user_fct, ncol=6) +
  labs (
    title = "Total carbs consumed trends by user",
    x = "Time",
    y = "Carbs"
  ) + 
  theme(
    axis.text.x = element_text(angle = 90)
  )
```


```{r}
meals.selected %>%
  mutate(eaten_at_date = as.Date(eaten_at, tz = "EST")) %>%
  filter(eaten_at_date <"2019-04-01" & !is.na(calories_eval)) %>%
  group_by(user_fct, eaten_at_date) %>%
  summarise(
    sum_carbs = sum(carbs_eval), 
    sum_protein = sum(protein_eval), 
    sum_fat = sum(fat_eval), 
    sum_fiber = sum(fiber_eval),
    total = sum_carbs+sum_protein+sum_fat+sum_fiber,
    Carbs=sum_carbs/total,
    Protein=sum_protein/total,
    Fat=sum_fat/total,
    Fiber=sum_fiber/total
  ) %>%
  gather(key = "Nutrient", value = "Proportion", Carbs, Protein, Fat, Fiber) %>%
  #print_df() %>%
  ggplot()+
  geom_point(aes(x=eaten_at_date, y=Proportion, color = Nutrient), size=0.1) +
  geom_smooth(aes(x=eaten_at_date, y=Proportion, color = Nutrient), method = "lm") +
  facet_wrap(~user_fct, ncol=6) +
  labs (
    title = "Nutrient proportions trends by user",
    x = "Time",
    y = "Proportion"
  ) + 
  guides(color = guide_legend(nrow = 1), fill = guide_legend(nrow = 1)) +
  theme(
    axis.text.x = element_text(angle = 90),
    legend.title = element_blank(), 
    legend.position="bottom",
  )
```



```{r}
bg_readings.all %>%
  filter(user_fct %in% selected_user_ids) %>%
  mutate(hour = format(as.POSIXct(time, format="%H:%M:%S"),"%H")) %>%
  group_by(user_fct, hour) %>%
  summarise(avg_bg = mean(bg)) %>%
  ggplot(aes(x=hour, y=avg_bg))+
  geom_point(size=0.3)+
  scale_x_discrete(breaks=c(0, 6, 12, 18, 24)) + 
  facet_wrap(~user_fct, ncol=6) +
  labs(
    title = "Average BG Level by Hour",
    x="Hour", 
    y="Average BG Level"
  ) +
  theme(
     axis.text.x = element_text(angle = 90),
  )
```

Have the ability of users to estimate nutrients change overtime with the use of the app?

```{r}
meals.imputed %>%
  filter(!is.na(calories_eval) & !is.na(calories_user)) %>%
  mutate(
    calories_diff = calories_user - calories_eval,
    carbs_diff = carbs_user - carbs_eval,
    protein_diff = protein_user - protein_eval,
    fat_diff = fat_user - fat_eval,
    fiber_diff = fiber_user - fiber_eval
  ) %>%
  mutate(eaten_at_date = as.Date(eaten_at, tz = "EST")) %>%
  filter(eaten_at_date <"2019-04-01") %>%
  group_by(user_fct, eaten_at_date) %>%
  summarise(carbs_diff = mean(carbs_diff)) %>%
  ggplot(aes(x=eaten_at_date, y=carbs_diff))+
  geom_point(size=0.1) +
  geom_smooth(method = "lm", na.rm=TRUE, size=0.5) +
  facet_wrap(~user_fct, ncol=6, scales = "free_y") +
  labs (
    title = "Trend of carbs estimation error",
    x = "Time",
    y = "Carbs user estimation error"
  ) + 
  theme(
    axis.text.x = element_text(angle = 90)
  )

```

```{r}
meals.imputed %>%
  filter(!is.na(calories_eval) & !is.na(calories_user)) %>%
  mutate(
    calories_diff = calories_user - calories_eval,
    carbs_diff = carbs_user - carbs_eval,
    protein_diff = protein_user - protein_eval,
    fat_diff = fat_user - fat_eval,
    fiber_diff = fiber_user - fiber_eval
  ) %>%
  gather(key = "type", value = "diff", carbs_diff, protein_diff, fat_diff) %>%
  ggplot() +
  geom_boxplot(aes(x = type, y = diff)) +
  coord_flip() +
  facet_wrap(~user_fct, ncol = 6)
```

## Sleep

```{r}
sleep = read_csv(file.path(raw_data_path, "1983", "sleep.csv"), col_types = cols()) %>%
  mutate(time = as.timestamp(time))
```


Questions:
* does the amount of sleep impact the first BG reading in the day (pre-breakfast, important!) (the very first pre-breakfst BG reading is considered fasting and is a true-ish representation of glycemic control (how bad is your diabetis)).
* Does amount (or quality of sleep) have an impoact on pre-breakfast BG (<- new measure). 
* Urban mith - the quality/amount of sleep may change how people respond to meals, e.g. for the same nutritional breakdown of a meal BG_impoact will be different.

* Fat and protein mediate the impact of carbohidrate. How to show this (focus on the visualization, data may be lacking the signal)
