---
title: "DataProcessing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(forcats)
library(r2d3)
library(parcoords)
library(dplyr)
```

## Data Import


```{r}
raw_data_path = "~/Dropbox/MealyzerData"
mealyser_data_overview = read_csv(file.path(raw_data_path, "mealyzer_data_overview.csv"), col_types = cols())
```


### Selecting users for the analysis

```{r}
users = mealyser_data_overview %>%
  filter(complete_meals > 14)
print(users[["user_id"]])
```

```{r}
read_mealyzer_files <- function(data_dir, user_ids, file_name, add_user_id, col_types) {
  result <- c()
  for (user_id in user_ids) {
    
    file_path = file.path(data_dir, as.character(user_id), file_name)
    df = read_csv(file_path, col_types = col_types)
    if (add_user_id) {
      df = cbind(user_id = user_id, df)
    }
    if (length(result) == 0) {
      result <- df
    } else {
      result <- rbind(result, df)
    }
  }
  return(result)
}

as.timestamp <- function(minutes_since, anchor_date = '2019-01-01 00:00:00 EST') {
  anchor_time = as.POSIXct(anchor_date)
  return(anchor_time + minutes_since * 60)
}

as.meals_kind <- function(values) {
  return(fct_relevel(values, c("breakfast", "morning_snack", "lunch", "afternoon_snack", "dinner", "after_dinner_snack")))
}

print_df <- function(df, ...) {
  print(df %>% select(...))
  return(df)
}

```

```{r}
bg_readings = read_mealyzer_files(raw_data_path, users[["user_id"]], "bg_labeled.csv", add_user_id = TRUE, col_types = cols(
  time = col_double(),
  bg = col_double(),
  meals_kind = col_character(),
  meal_id = col_double(),
  readings_kind = col_character()
)) %>%
  mutate(
    user_id = as.character(user_id),
    time = as.timestamp(time), 
    meals_kind = as.meals_kind(meals_kind),
    readings_kind = fct_relevel(readings_kind, c("premeal", "postmeal"))
  )
head(bg_readings)
```
```{r}
# No NA's in the bg_readings file
extracat::visna(bg_readings)
```


```{r}
meals_data = read_mealyzer_files(raw_data_path, users[["user_id"]], "mealyzer_data.csv", add_user_id = FALSE, col_types = cols()) %>%
  mutate(
    user_id = as.character(user_id),
    eaten_at = as.timestamp(eaten_at),
    kind = as.meals_kind(kind),
    premeal_bg_time = as.timestamp(premeal_bg_time),
    postmeal_bg_time = as.timestamp(postmeal_bg_time),
    premeal_bg_delay_minutes = as.double(difftime(eaten_at, premeal_bg_time, units = "mins")),
    postmeal_bg_delay_minutes = as.double(difftime(postmeal_bg_time, eaten_at, units = "mins")),
    bg_impact = postmeal_bg - premeal_bg,
    bg_impact_slope = bg_impact / (premeal_bg_delay_minutes + postmeal_bg_delay_minutes)
  )
head(meals_data)
```
```{r}
extracat::visna(meals_data)
```

### Dataset highlevel view

```{r fig.width=18}
meals_data %>%
  mutate(eaten_at_hour = hour(eaten_at)) %>%
  ggplot() +
  geom_histogram(aes(x=eaten_at_hour), bins = 24) +
  facet_wrap(~user_id, ncol = 7) + 
  labs(
    title = "Distribution of hours when users had a meal"
  ) + 
  theme(
    strip.text = element_text(size=25)
  )
```

```{r fig.width=18}
premeal_delay_buckets = c("< 15 min", "10 - 60 min", "> 60 min")
meals_data %>%
  #filter(user_id == 2392) %>%
  #filter(premeal_bg_delay_minutes <= 300 & premeal_bg_delay_minutes >= 0) %>%
  #mutate(premeal_bg_delay = ifelse(premeal_bg_delay_minutes < 15, premeal_delay_buckets[1], ifelse(premeal_bg_delay_minutes < 60, premeal_delay_buckets[2], premeal_delay_buckets[3]))) %>%
  #mutate(premeal_bg_delay = fct_relevel(premeal_bg_delay, premeal_delay_buckets)) %>%
  #print_df(user_id, premeal_bg_time, eaten_at, premeal_bg_delay_minutes) %>%
  ggplot() +
  #geom_bar(aes(x=premeal_bg_delay, fill=premeal_bg_delay)) +
  geom_histogram(aes(x=premeal_bg_delay_minutes), binwidth = 30) +
  scale_x_continuous(breaks=c(0, 60, 120, 180, 240, 300), labels = c("0", "1h", "2h", "3h", "4h", "5h"), limits = c(0, 300)) + 
  facet_wrap(~user_id, ncol = 7, scales = "free_y") + 
  labs(
    title = "Delay of premeal BG recordings, bucketed"
  ) + 
  theme(
    legend.position="top",
    #axis.title.x = element_blank(),
    #axis.text.x = element_blank(),
    #axis.ticks.x = element_blank()
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.text = element_text(size=25),
  )
```

```{r fig.width=18}
postmeal_delay_buckets = c("< 1 hour", "1 - 2 hours", "> 2 hours")
meals_data %>%
  #mutate(postmeal_bg_delay = ifelse(postmeal_bg_delay_minutes < 60, postmeal_delay_buckets[1], ifelse(postmeal_bg_delay_minutes < 120, postmeal_delay_buckets[2], postmeal_delay_buckets[3]))) %>%
  #mutate(postmeal_bg_delay = fct_relevel(postmeal_bg_delay, postmeal_delay_buckets)) %>%
  #print_df(user_id, eaten_at, postmeal_bg_time, postmeal_bg_delay_minutes) %>%
  ggplot() +
  #geom_bar(aes(x=postmeal_bg_delay, fill=postmeal_bg_delay)) +
  geom_histogram(aes(x=postmeal_bg_delay_minutes), binwidth = 30) +
  #geom_density(aes(x = postmeal_bg_delay_minutes)) +
  facet_wrap(~user_id, ncol = 7) + 
  scale_x_continuous(breaks=c(0, 60, 120, 180, 240, 300), labels = c("0", "1h", "2h", "3h", "4h", "5h"), limits = c(0, 300)) + 
  labs(
    title = "Delay of postmeal BG recordings, bucketed"
  ) + 
  theme(
    legend.position="top",
    axis.title.x = element_blank(),
    #axis.text.x = element_blank(),
    #axis.ticks.x = element_blank()
    #axis.title.y = element_blank(),
    #axis.text.y = element_blank(),
    #axis.ticks.y = element_blank()
    strip.text = element_text(size=25),
  )
```

```{r fig.height=12, fig.width=6}
meals_data %>%
  ggplot() +
  geom_boxplot(aes(x = user_id, y = premeal_bg_delay_minutes), varwidth = TRUE) + 
  scale_y_continuous(breaks=c(0, 15, 30, 45, 60), labels = c("0", "15min", "30min", "45min", "1h"), limits = c(0, 60)) + 
  coord_flip() +
  labs( title = "" )
```

```{r fig.height=12, fig.width=6}
meals_data %>%
  ggplot() +
  geom_boxplot(aes(x = user_id, y = postmeal_bg_delay_minutes), varwidth = TRUE) + 
  scale_y_continuous(breaks=c(0, 60, 120, 180, 240), labels = c("0", "1h", "2h", "3h", "4h"), limits = c(0, 1000)) + 
  coord_flip() +
  labs( title = "" )
```



## Per user analysis

```{r}
focus_user_id = "2475"
```

```{r fig.width=18, fig.height=18}
library(lattice) #sploms

meals_data %>%
  filter(user_id == focus_user_id) %>%
  mutate(
    protein_fat=eval_proportion_protein-eval_proportion_fat, 
    protein_carb=eval_proportion_protein-eval_proportion_carbs, 
    carb_fat=eval_proportion_carbs-eval_proportion_fat) %>%
  select(calories_eval, carbs_eval, fat_eval, protein_eval, fiber_eval, premeal_bg, postmeal_bg, bg_impact, protein_fat, protein_carb, carb_fat) %>%
  splom()
```

```{r fig.width=18, fig.height=18}
library(lattice) #sploms

meals_data %>%
  filter(user_id == focus_user_id) %>%
  mutate(
    protein_fat_log_ratio=log(eval_proportion_protein/(eval_proportion_fat+0.0000001)), 
    protein_carb_log_ratio=log(eval_proportion_protein/(eval_proportion_carbs+0.0000001)), 
    carb_fat_log_ratio=log(eval_proportion_carbs/(eval_proportion_fat+0.0000001))
  ) %>%
  select(eval_calories_computed, eval_proportion_carbs, eval_proportion_fat, eval_proportion_protein, premeal_bg, postmeal_bg, bg_impact, protein_fat_log_ratio, protein_carb_log_ratio, carb_fat_log_ratio) %>%
  splom()
```
Talking points:

* Correlation between calories_eval and all carbs|fat|protein, but much smaller correlation to fiber
* premeal_bg and postmeal_bg - positively correlated
* negative correlation between eval_proportion_carbs & eval_proportion_fat
* interactions between ..._log_ratio varables split the meals into 2 or 3 clusters


### Time effect

```{r fig.width=15}
meals_data %>% 
  #filter(user_id == focus_user_id) %>%
  ggplot() +
  facet_wrap(~user_id, ncol = 7) + 
  geom_segment(aes(x = -premeal_bg_delay_minutes, xend = postmeal_bg_delay_minutes, y=premeal_bg, yend=postmeal_bg), alpha = 0.5) + 
  scale_x_continuous(breaks=c(-15, 0, 60, 120, 180, 240), labels = c("-15min", "0", "1h", "2h", "3h", "4h"), limits = c(-15, 180)) + 
  theme(
    strip.text = element_text(size=25)
  )

```

```{r fig.height=5, fig.width=9}
display_meals <- function(df, ncol = 6, nrow = 3, rotate = 0) {
  meals = df[1:(ncol*nrow), ] %>% select(user_id, meal_id)
  filenames = paste0(normalizePath("~/Dropbox/MealyzerData"), "/", meals[["user_id"]], "/images/medium/", meals[["meal_id"]], ".jpg")
  #print(filenames)
  data_uris = sapply(filenames, knitr::image_uri)
  images = tibble(src = data_uris)
  r2d3(data=images, script = "display_meals.js", container = "div", options=list(ncol=ncol, nrow=nrow, rotate=rotate))
}

meals_data %>%
  filter(user_id == focus_user_id) %>%
  arrange(desc(calories_eval)) %>%
  #print_df(bg_impact, title, ingredients, calories_eval) %>%
  display_meals(ncol = 6, nrow = 3, rotate = -90)
```

Talking point:
* there are a few testing images users made, and they will need to be found and ignored for future analysis.

```{r}
meals_data %>%
  filter(user_id == focus_user_id) %>%
  arrange(calories_eval) %>%
  #print_df(bg_impact, title, ingredients, calories_eval) %>%
  display_meals(ncol = 6, nrow = 3, rotate = -90)
```


## Meal nutrients evaluator vs user

```{r}
meals_data %>%
  mutate(calories_diff = calories_user - calories_eval) %>%
  filter(!is.na(calories_diff)) %>%
  ggplot() +
  geom_histogram(aes(x = calories_diff), binwidth = 100)
```

```{r }
meals_data %>%
  mutate(
    calories_diff = calories_user - calories_eval,
    carbs_diff = carbs_user - carbs_eval,
    protein_diff = protein_user - protein_eval,
    fat_diff = fat_user - fat_eval,
    fiber_diff = fiber_user - fiber_eval
  ) %>%
  filter(!is.na(calories_diff) & !is.na(carbs_diff) & !is.na(protein_diff) & !is.na(fat_diff) & !is.na(fiber_diff)) %>%
  gather(key = "type", value = "diff", carbs_diff, protein_diff, fat_diff, fiber_diff) %>%
  ggplot() +
  #facet_wrap(~user_id, ncol = 6, scales = "free_x") + 
  geom_boxplot(aes(x = reorder(type, diff, median), y = diff)) +
  coord_flip()
```

Talking point:

* Slight under-counting of the calories, but not as much as I would expect. High density at 0, suggesting lots of meal the evaluator and user entered exactly the same number of calories, possibly because of 0 calory meal (e.g. water) or known calory meal (e.g. protein bar)
* Systematic under-counting of carbs, slight over-counting of fiber

## Overtime trends

```{r fig.width=5}

meals_data %>%
  ggplot() +
  geom_point(aes(x = eaten_at, y = calories_eval)) +
  facet_wrap(~user_id, ncol = 7, scales = "free_x") 
```  

Have the ability of users to estimate nutrients change overtime with the use of the app?


### Plotting activeness vs. bg level

```{r}
bg_readings %>%
  group_by(meal_id) %>%
  mutate(change_in_bg = c(0,diff(bg)/first(bg))) %>%
  select(user_id, change_in_bg) %>%
  filter(change_in_bg != 0) %>%
  group_by(user_id) %>%
  summarise(n = n(), avg_change_in_bg = mean(change_in_bg))  %>%
  mutate(activeness = ifelse(n > median(n), "high","low"))  %>%
  ggplot(aes(x=seq_along(avg_change_in_bg), y=avg_change_in_bg)) +
  geom_point(aes(colour=activeness)) 
```


### Plotting pre/post meal bg ,faceting on user_id

```{r}
bg_readings %>%
  group_by(meal_id) %>%
  mutate(premeal = first(bg), postmeal =last(bg)) %>%
  select(user_id, meals_kind, premeal, postmeal) %>%
  distinct() %>%
  ggplot(aes(x=premeal, y=postmeal)) +
  geom_point(size = 0.2, aes(colour=meals_kind)) +
  facet_wrap(~user_id, ncol = 7)
```


```{r}
meals_data %>%
  mutate(eaten_at_date = as.Date(eaten_at, tz = "EST")) %>%
  filter(eaten_at_date <"2019-04-01") %>%
  group_by(user_id, eaten_at_date) %>%
  summarise(sum_by_day = sum(calories_eval)) %>%
  ggplot(aes(x=eaten_at_date, y=sum_by_day))+
  geom_point(size=0.3)+
  geom_smooth(method = "lm")+
  facet_wrap(~user_id, ncol=7) +
  labs (
    title = "Calories trends by user"
  )
```

```{r}
meals_data %>%
  mutate(eaten_at_date = as.Date(eaten_at, tz = "EST")) %>%
  filter(eaten_at_date <"2019-04-01") %>%
  group_by(user_id, eaten_at_date) %>%
  summarise(sum_by_day = sum(carbs_eval)) %>%
  ggplot(aes(x=eaten_at_date, y=sum_by_day))+
  geom_point(size=0.3)+
  geom_smooth(method = "lm")+
  facet_wrap(~user_id, ncol=7) +
  labs (
    title = "Carbs trends by user"
  )
```


```{r fig.width=4}
dplyr::summarise(group_by(meals_data, user_id, kind), freq = n()) %>%
  group_by(user_id) %>%
  mutate(prop = freq/sum(freq)) %>%
  ungroup() %>%
  ggplot(aes(x = user_id, y = prop, fill = forcats::fct_rev(kind))) +
      geom_bar(stat = "identity") + 
      coord_flip() + 
      labs(x = "User ID",
           y = "Proportion",
           title = "Proportion of All Meals by Meal Kind") + 
           theme(legend.title = element_blank(), legend.position="bottom") +
  guides(fill = guide_legend(nrow = 1, reverse=T))
```
```{r fig.width=4}
meals_data %>%
  select(user_id,carbs_eval, protein_eval,fat_eval,fiber_eval) %>%
  gather(key = "variable", value, -user_id) %>%
  group_by(user_id, variable) %>%
  summarise(total=sum(value, na.rm = T)) %>%
  group_by(user_id) %>%
  mutate(prop = total/sum(total)) %>%
  ungroup() %>%
  ggplot(aes(x = user_id, y = prop, fill = forcats::fct_rev(variable))) +
      geom_bar(stat = "identity") + 
      coord_flip() + 
      labs(x = "User ID",
           y = "Proportion",
           title = "Proportion of All Meals by Nutrient") + 
           theme(legend.title = element_blank(), legend.position="bottom") +
  guides(fill = guide_legend(nrow = 1, reverse=T))
```

```{r}
parcoords(
  meals_data %>% 
    select(kind,carbs_eval, protein_eval,fat_eval,fiber_eval,calories_eval, bg_impact)
  , rownames = F # turn off rownames from the data.frame
  , brushMode = "1D-axes"
  , reorderable = T
  , queue = T,
  alpha = 0.05,
  color = list(
      colorBy = "kind"
      , colorScale = "scaleOrdinal"
      , colorScheme = "schemeCategory10"
    )
  , withD3 = TRUE
)
```
```{r}
parcoords(
  meals_data %>% 
    select(user_id, kind, bg_impact, calories_eval, carbs_eval, protein_eval,fat_eval,fiber_eval) %>%
    gather(key = "variable", value, -user_id,-kind) %>%
    group_by(user_id, kind, variable) %>%
    summarise(avg_eval=mean(value, na.rm = T)) %>%
    spread(variable, avg_eval)
  , rownames = F # turn off rownames from the data.frame
  , brushMode = "1D-axes"
  , reorderable = T
  , queue = T,
  alpha = 0.3,
  color = list(
      colorBy = "kind"
      , colorScale = "scaleOrdinal"
      , colorScheme = "schemeCategory10"
    )
  , withD3 = TRUE
)
```

```{r fig.height=4,fig.width=6}
meals_data %>%
  select(user_id, kind,carbs_eval, protein_eval,fat_eval,fiber_eval,calories_eval, bg_impact) %>%
  mutate_if(is.numeric, scale) %>%
  mutate(ID = 1:3303) %>%
  gather(key = "variable", value, -user_id,-kind, -ID) %>% 
  ggplot(aes(x=variable,y=value,colour=kind, group=ID))+
  geom_line(alpha=0.5)+facet_wrap(~user_id,ncol=7)+
  theme(
    legend.position="top",
    axis.text=element_text(size=5, angle = 90)
  )
```

```{r}
meals_data %>%
  select(user_id, kind,carbs_eval, protein_eval,fat_eval,fiber_eval, bg_impact) %>%
  #mutate(ID = 1:3303) %>%
  gather(key = "variable", value, -user_id,-kind) %>%
  group_by(user_id, kind, variable) %>%
  summarise(avg_eval=mean(value, na.rm = T)) %>%
  ungroup() %>%
  mutate_if(is.numeric, scale) %>%
  #gather(key = "variable", value, -user_id,-kind, -ID, -avg_eval) %>% 
  ggplot(aes(x=variable,y=avg_eval,group=1))+
  geom_line(alpha=0.5)+
  coord_flip() +
  facet_wrap(~user_id,ncol=7)
```

```{r}
focus_user_id = "1983"
sleep = read_csv(file.path(raw_data_path, focus_user_id,"sleep.csv"), col_types = cols()) %>%
  mutate(time = as.timestamp(time))
```

```{r}
sleep %>%
  ggplot() +
  geom_point(aes(x=time, y= seconds,color=factor(level)), size=0.5)
```
```{r}
sleep %>%
 filter(level %in% c("deep","light","rem","wake")) %>%
  ggplot(aes(x = seconds)) +
  geom_histogram(alpha=.7, color="black", fill="steelblue2", bins = 32) +
  facet_wrap( ~level) +
  xlab("Time Frame in Seconds") + ylab("Count") + 
  ggtitle("Histogram of Time Frames By Sleep Stage") +
  theme(plot.title = element_text(face = "bold"))
```

```{r}
library("ggridges")
sleep %>%
  filter(level %in% c("deep","light","rem","wake")) %>%
  ggplot(aes(x = seconds, y = level, fill = level)) +
  geom_density_ridges_gradient(scale = 2) + theme_ridges()+
  scale_y_discrete(expand = c(0.3, 0)) +
  scale_x_continuous(expand = c(0.01, 0))+
  labs(x="Time Frames", y="Sleep Stage")+
  ggtitle("Density Curves of Time Frames by Sleep Stage")+
  theme(plot.title = element_text(hjust = 0.6)) +
  guides(fill = guide_legend(reverse=T))

```

```{r}
bg_readings %>%
  filter(time <"2019-05-01 00:00:00") %>%
  ggplot(aes(x = time, y = bg)) +
  geom_line() +
  facet_wrap(~user_id,ncol=7) +
  theme(axis.text=element_text(size=5, angle = 90))
```

```{r}
bg_readings %>%
  #mutate(time = round_date(time, '1 hours')) %>%
  mutate(time = format(as.POSIXct(time,format="%H:%M:%S"),"%H")) %>%
  group_by(user_id, time) %>%
  summarise(avg_bg = mean(bg)) %>%
  ggplot(aes(x=time, y=avg_bg))+
  geom_point(size=0.3)+
  facet_wrap(~user_id, ncol=7) +
  labs(x="Hour", y="Average BG Level")+
  ggtitle("Average BG Level by Hour") +
  theme(plot.title = element_text(face = "bold")) +
  theme(axis.text=element_text(size=5, angle = 90))
```

```{r}
distance = read.csv("~/Dropbox/MealyzerData/1983/activities-distance_overview.csv", header=T)
calories = read.csv("~/Dropbox/MealyzerData/1983/activities-calories_overview.csv", header=T)
steps = read.csv("~/Dropbox/MealyzerData/1983/activities-steps_overview.csv", header=T)

activities <- merge(steps, merge(distance, calories, by = "date", all = TRUE), by = "date", all = TRUE)
colnames(activities) <- c("date","steps","distance","calories")

# date vs. calories,distance,steps for user #1983
activities %>% 
  mutate(date = date(as.timestamp(date))) %>%
  slice(1:133) %>% # weird data after this point
  gather(key = key, value = value, -date) %>%
  ggplot(aes(date, value, col=key)) + geom_line()

# scale the above data
activities %>% 
  mutate(date = date(as.timestamp(date))) %>%
  slice(1:133) %>% # weird data after this point
  gather(key = key, value = value, -date) %>%
  group_by(key) %>%
  mutate(index = round(100*value/value[1], 2)) %>% 
  ungroup() %>%
  ggplot(aes(date, index, col=key)) + geom_line()

# once scaled, we see that the data for distance and steps completely overlaps, suggesting that one is a linear function of another.  
```

```{r}
user1983 = bg_readings %>%
  mutate(date=date(time)) %>%
  group_by(user_id, date) %>%
  mutate(avg_bg=mean(bg)) %>%
  select(user_id,date,avg_bg) %>%
  unique() %>%
  filter(user_id==1983) %>%
  slice(1:72) # weird data after this point

user1983 %>%
  ggplot(aes(date, avg_bg)) + geom_line()

# plot avg_bg vs. calories for user #1983
activities %>% 
  mutate(date = date(as.timestamp(date))) %>%
  full_join(user1983, by="date") %>%  #or use merge?
  select(-user_id, -distance, -steps) %>%
  filter(date < "2019-04-01") %>%
  gather(key = key, value = value, -date) %>%
  group_by(key) %>%
  mutate(index = round(100*value/value[1], 2)) %>% 
  ungroup() %>%
  ggplot(aes(date, index, col=key)) + geom_line()

# We see that there are missing values for average BG, which is normal since bg level and calories burnt are measured by two different devices/apps.
# Based on the data we have here, we see that avg_bg has a nagative relationship with calories burnt, where higher avg_bg usually is associated with lower calories burnt that day.
```



```{r}
#user 1983's avg_bg vs. minutesasleep by day

sleep_overview = read.csv("~/Dropbox/MealyzerData/1983/sleep_overview.csv")
sleep_overview %>%
  mutate(date=date(as.timestamp(dateOfSleep))) %>%
  arrange(date) %>% # to perform correct scaling, must have date in ascending order
  group_by(date) %>%
  mutate(minutesAsleep = sum(minutesAsleep)) %>%
  full_join(user1983, by="date") %>% #merge?
  select(date, minutesAsleep, avg_bg) %>%
  filter(date < "2019-04-01") %>%
  gather(key = key, value = value, -date) %>%
  group_by(key) %>%
  mutate(index = round(100*value/value[1], 2)) %>% 
  ungroup() %>%
  ggplot(aes(date, index, col=key)) + geom_line()

```

```{r}
# user 1983's parcoords 
parcoords(
  sleep_overview %>% 
    select(-minutesToFallAsleep) %>%
    mutate(date=date(as.timestamp(dateOfSleep))) %>%
    group_by(date) %>%
    mutate(minutesAsleep = sum(minutesAsleep)) %>%
    merge(user1983, by="date") %>%
    ungroup() %>%
    select(-minutesAfterWakeup, -date, - user_id) 
  , rownames = F # turn off rownames from the data.frame
  , brushMode = "1D-axes"
  , reorderable = T
  , queue = T,
  alpha = 0.5,
  color = list(
      colorBy = "kind"
      , colorScale = "scaleOrdinal"
      , colorScheme = "schemeCategory10"
    )
  , withD3 = TRUE
)
# arrange in this order:
# dateOfSleep, endTime, startTime, minutesAwake, minutesAsleep, efficiency, ang_bg, duration, timeInBed
# Looks like dateOfSleep, endTime, startTime are the same, so is duration, timeInBed
# avg_bg is correlated with the duration - which may be a good talking point
```


